Citation Analysis of Computer Science

### 2019-11-16 Scientometrics/comp.tex
updates

%
Computer science, and its applications, has experienced dramatic growth and diversification over the last twenty years. As noted in a 2017 US National Academies Report~\cite{nas_2017}, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission". Additionally, the collective influence of the Internet of Things, `big' data, accessible cloud computing, and advances in artificial intelligence have been postulated as a recent driver for growth and evolution~\cite{siebel2019_digital}.

In a historical study of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles), Salton and Bergmark described the global structure of computer science as consisting of three main subareas (i) theoretical foundations, e.g., theory of computation (ii) hardware and computer systems, e.g., architecture (iii) software, e.g, programming systems~\cite{salton_citation_1979}.  Related areas discussed were  (a) mathematics of computing, e.g., numerical analysis (b) special software topics, e.g., operating systems (c) data management and database systems (d) methodologies valid for multiple applications, e.g., algebraic manipulation (e) computer applications, e.g., computer graphics (f) nontechnical aspects, e.g., computer education. In presenting this work, Salton and Bergmark made the key point that the scientific literature serves a rich source of information to study the structure and historical development of a field and reference relevant work in bibliometrics.

Progressing beyond the triad of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS) published by the Association for Computing Machinery (ACM) now consists of 13 top-level areas. https://dl.acm.org/ccs/ccs.cfm \emph{need to figure out to put this in BibTeX} that reflect a more current view of the field. This classification addresses relationships with other fields under Applied Computing. However, an easy way to map scientific publications, especially interdisciplinary articles or those from adjacent fields, to the CCS classification does not exist. Other classification systems exist such as the All Science Journal Classification (ASJC) developed and maintained by Scopus. the Web of Science classification of research areas, and the NSF classification system. All three rely on applying one or more journal-derived labels to articles. Limitations of journal based classification have stimulated efforts to use article level clusters using citations and text similarity for links. \emph{cite Boyack, Small, Waltman, etc,}. These limitations are well known~\emph{wang and waltman; lariviere paper} and efforts have been made to instead build article level clusters using citations or other criteria for assembly .
\emph{https://www.nsf.gov/statistics/nsf13327/pdf/tabb1.pdf}, WoS \\ 
\emph{https://bit.ly/2NBRqaH}, and the  
\\ \emph{insert url} are based on journals that rely on assigning journal-level labels to articles.

In this article, we explore the structure of computer science through matching records from two major data sources. (i) the dblp bibliography- a reference bibliography for computer science\emph{cite} (ii) the Scopus bibliography, which covers the physical science, life sciences, social sciences, and health sciences. Using direct citations and co-citations as links, we then construct clusters of articles that are drawn from journals and conference proceedings  and reconcile them with the All Science Journal Classification (ASJC) developed and maintained by Scopus.

@@

\emph{Data} The dblp computer science bibliography consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz from  `https://dblp.uni-trier.de/xml/'. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of the construction of a larger data platform for research evaluation~\cite{Korobskiy_2019}. Records in the dblp dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications with Scopus identifiers. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par

>>>>

Computer science, and its applications, has experienced dramatic growth and diversification over the last twenty years. As noted in a 2017 US National Academies Report~\cite{nas_2017}, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission". Additionally, the collective influence of the Internet of Things, `big' data, accessible cloud computing, and advances in artificial intelligence have been postulated as a recent driver for growth and evolution~\cite{siebel2019_digital}.

In a historical study of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles), Salton and Bergmark described the global structure of computer science as consisting of three main subareas (i) theoretical foundations, e.g., theory of computation (ii) hardware and computer systems, e.g., architecture (iii) software, e.g, programming systems~\cite{salton_citation_1979}.  Related areas discussed were  (a) mathematics of computing, e.g., numerical analysis (b) special software topics, e.g., operating systems (c) data management and database systems (d) methodologies valid for multiple applications, e.g., algebraic manipulation (e) computer applications, e.g., computer graphics (f) nontechnical aspects, e.g., computer education. In presenting this work, Salton and Bergmark made the key point that the scientific literature serves a rich source of information to study the structure and historical development of a field and cite similar ideas.

Progressing beyond this historical triad consisting of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS)~\cite{acm_ref} published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field. This classification also addresses relationships with other fields under Applied Computing. However, an easy way to map scientific publications, especially interdisciplinary articles or those from adjacent fields, to the CCS classification does not exist. Other classification systems exist such as the All Science Journal Classification (ASJC) developed and maintained by Scopus. the Web of Science research and categorical classifications, and the NSF classification system. All three rely on applying one or more journal-derived labels to articles. Classification systems that use articles directly have also been developed~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, which escape some of the criticisms of journal-based classification~\cite{wang_large-scale_2016,shu_comparing_2019}.

In this article, we explore the structure of computer science through matching records from two major data sources. (i) the dblp bibliography- a reference bibliography for computer science~\cite{dblp_ref} (ii) the Scopus bibliography, which covers the physical science, life sciences, social sciences, and health sciences. Using direct citations and co-citations as links, we then construct clusters of articles that are drawn from journals and conference proceedings  and reconcile them with the All Science Journal Classification (ASJC) developed and maintained by Scopus.

@@

\emph{Data} A stable release of the dblp computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of the construction of a larger data platform for research evaluation~\cite{Korobskiy_2019}. Records in the dblp dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications with Scopus identifiers. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par
